### 2016-05-09 capability-machine.tex
Small latex fix.
Added O relation and fixed a couple of typos.
still work in progress
Updated the ticket dispenser code to use a capability stored on the heap
Update some notations wrt updating heap/registers
Updated semantics and added the remaining instructions needed for example.

\newcommand{\update}[2]{[#1 \mapsto #2]}%TODO change this to nicer notation \Phi[heap.a \mapsto w]
\newcommand{\sem}[1]{\left\llbracket #1 \right\rrbracket}

@@

@@

%Configuration end states
\newcommand{\failed}{\mathit{failed}}
\newcommand{\halted}{\mathit{halted}}

%Functions
\newcommand{\plainfun}[1]{\mathit{#1}}
\newcommand{\decode}{\plainfun{decode}}
\newcommand{\encode}{\plainfun{encode}}
\newcommand{\encodePerm}{\plainfun{encodePerm}}
\newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}}
\newcommand{\executeAllowed}[1]{\plainfun{executeAllowed}(#1)}
\newcommand{\nonZero}[1]{\plainfun{nonZero}(#1)}
\newcommand{\readAllowed}[1]{\plainfun{readAllowed}(#1)}
\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}(#1)}

%World operations
\newcommand{\future}{\mathbin{\sqsupseteq}}
\newcommand{\heapSat}[3][\Phi]{#1 :_{#2} #3}

@@

%LR
\newcommand{\intr}[2]{\mathcal{#1}\sem{#2}}
\newcommand{\valueintr}[1]{\intr{V}{#1}}
\newcommand{\exprintr}[1]{\intr{E}{#1}}
\newcommand{\contintr}[1]{\intr{K}{#1}}
\newcommand{\stdvr}{\valueintr{\asmType}}
\newcommand{\stder}{\exprintr{\asmType}}
\newcommand{\stdkr}{\contintr{\asmType}}
\newcommand{\npair}[2][n]{\left(#1,#2 \right)}

@@

\begin{document}
$\RegName$ contains $\pcreg$, but is otherwise some undefined, finite set.
\begin{align*}
\Addrs &::= \nats & & &
\Words &::= \Caps + \ints \\
\Regs  &::= \RegName \rightarrow \Words & & &
\Heaps &::= \Addrs \rightarrow \Words \\
\Perms &::= \{ \noperm, \readonly, \readwrite, \exec, \entry, \rwx\} & & &
\Mems  &::= \Regs \times \Heaps \\
\Caps  &::= \Perms \times \Addrs \times \Addrs \times \Addrs & & &
\Confs &::= \Mems + \{\failed, \halted\}
\end{align*}

@@

  \path[every node/.style={font=\sffamily\small}]
    (1) edge (2)
    (2) edge (3)
    (3) edge (6)
    (1) edge (4)
    (4) edge (5)
    (5) edge (6);
\end{tikzpicture}
\caption{Permission hierarchy}
\label{fig:perm-hier}
\end{figure}
Notation:
$$\begin{array}{rcl}
i       &\in& \Instrs \\
r       &\in& \RegName\\
\mem    &::=& (\reg,\heap)\\
\pc     &\in& \Caps \\
\pcreg  &\in& \RegName \\
\Phi    &::=& \mem \in \Confs\\
\addr   &\in& \Addrs\\
\perm   &\in& \Perms\\
(\perm,\start,\addrend,\addr) &\in& \Caps \\
\end{array}$$
Further definitions:
$$\begin{array}{rcl}
\lv    &::=& \refreg{r} \\
\hv    &::=& \refheap{r}\\
\rv    &::=& n \mid \lv \\
i      &::=& \fail \mid \halt \mid 
             \jmp{\lv} \mid \jnz{\lv}{\rv} \mid
             \isptr{\lv}{\rv} \mid \setptr{\lv}{\rv} \mid \\
       &   & \lea{\lv}{\rv} \mid\move{\lv}{\rv} \mid \load{\lv}{\hv} \mid \store{\hv}{\rv} \mid  \\
       &   & \restrict{\lv}{\rv}{\rv} \mid \subseg{\lv}{\rv}{\rv} \mid \plus{\lv}{\rv}{\rv}
\end{array}$$

\subsection*{Semantics}
\begin{align*}
\decode &:\ints \rightarrow \Instrs
\end{align*}
%TODO for some decode function

@@

\Phi & \rightarrow \sem{\decode(\Phi.\reg(\pcreg))}(\Phi) & & \text{if $\Phi.\reg(\pcreg) = (\exec{\_},{\_},{\_})$} \\

@@

@@

%TODO: Update semantics according to changes.

@@

  \sem{\halt}(\Phi)     & = \halted \\
  \sem{\jmp{\lv}}(\Phi) & = \begin{cases}
                            (\Phi.\reg\update{\pcreg}{\updatePcPerm{c}}) & \text{if }\Phi.reg(lv) = c \\
                                                                         & \text{  and }c=(\perm,\start,addrend,\addr)\\
                                                                         & \text{  and }\executeAllowed{\perm}\\

@@

  \sem{\jnz{\lv}{\rv}}(\Phi) & = \begin{cases}
                            \Phi.\reg\update{\pcreg}{\updatePcPerm{\var{c}}} &

@@

                              \text{if $\nonZero{\Phi.\reg(\rv)}$} \\ 
                              \text{  and $\Phi.reg(lv) = c$} \\
                              \text{  and $c=(\perm,\start,\addrend,\addr)$}\\
                              \text{  and $\executeAllowed{\perm}$}

@@

                            \Phi.\reg\update{\pcreg}{\Phi.\reg(\pcreg) + 1} & \text{if not $\nonZero{\Phi.\reg(\rv)}$}\\

@@

                                   \Phi.\reg\update{r_1}{\var{w}} &

@@

                                     \text{if }\Phi.\reg(r_2) = (\perm,\start,\addrend,\addr)\\
                                     \text{  and }\readAllowed{\perm} \text{ and } \var{w} = \Phi.\heap(\addr)

@@

                                   \Phi.\heap\update{\addr}{\var{w}} &

@@

                                     \text{if }\Phi.\reg(r_1) = (\perm,\start,\addrend,\addr)\\
                                     \text{  and }\writeAllowed{\perm} \text{ and } \var{w} = \Phi.\reg(r_2)

@@

 \sem{\move{\refreg{r_1}}{\rv}} & =

@@

                                   \Phi.\reg\update{r_1}{\Phi.\reg(\rv)} & \text{if $r_1 \neq \pcreg$} \\

@@

@@

\section{Examples}
\label{sec:examples}
\subsection{Ticket Dispenser}
\newcommand{\size}{\var{size}}
\label{sec:tick-disp}
\[
  \begin{array}{r l}
    \init: & \move{\refreg{r_1}}{\size} \\
           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
           & \lea{\refreg{r_0}}{3} \\
           & \jmp{\malloc} \\
           & \store{\refheap{r_1}}{(\encode(i_1))} \\
           & \lea{\refreg{r_1}}{1} \\
           & \store{\refheap{r_1}}{(\encode(i_2))} \\
           & \lea{\refreg{r_1}}{1} \\
           & \vdots \\
           & \store{\refheap{r_1}}{(\encode(i_6))} \\
           & \lea{\refreg{r_1}}{1} \\
           & \move{\refreg{r_0}}{\refreg{r_1}} \\
           & \lea{\refreg{r_0}}{1} \\
           & \store{\refheap{r_1}}{\refreg{r_0}} \\
           & \lea{\refreg{r_1}}{1} \\
           & \store{\refheap{r_1}}{0} \\
           & \lea{\refreg{r_1}}{-7} \\
           & \restrict{\refreg{r_1}}{(\encodePerm{(\entry)})} \\
           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
           & \lea{\refreg{r_0}}{4} \\
           & \restrict{\refreg{r_0}}{(\encodePerm{(\entry)})} \\
           & \jmp{\codelabel{adv}}
\]
Here $\size$ is 8. $i_1 \dots i_7$ refers to the instructions in the following program:
\[
  \begin{array}{r l}
    i_1 :& \load{\refreg{r_2}}{\size-1} \\
    i_2 :& \load{\refreg{r_1}}{\refheap{r_2}} \\
    i_3 :& \plus{\refreg{r_3}}{\refreg{r_1}}{2} \\
    i_4 :& \store{\refheap{r_2}}{\refreg{r_3}} \\
    i_5 :& \move{\refreg{r_2}}{0} \\
    i_6 :& \jmp{\refreg{r_0}}
\]
The heap layout for the above program can be seen in Figure~\ref{tab:tick-disp-heap}.
\begin{figure}[h]
  \centering
  \begin{tabular}{ | c | l | }
    \hline
    Addr. & Content \\ \hline
    1   &  $i_1$ \\ \hline
    2   &  $i_2$ \\ \hline
    3   &  $i_3$ \\ \hline
    4   &  $i_4$ \\ \hline
    5   &  $i_5$ \\ \hline
    6   &  $i_6$ \\ \hline
    7   & Cap. for addr. 8 \\ \hline
    8   & The counter \\ \hline  
  \end{tabular}
  \caption{The heap layout for the ticket dispenser.}
  \label{tab:tick-disp-heap}
\end{figure}

\section{Logical Relation}
\label{sec:logical-relation}
\subsection{Worlds}
\[
\Worlds \cong \frac{1}{2} (\RegionName \parfun (\Worlds \monfun \UPred{\Addrs \parfun \Worlds}))
\]

\[
\stdvr \; : \;  \Worlds \monfun \UPred{\Words}
\]
\begin{align*}
  \stdvr(W) \defeq & \{ \npair{i} \mid i \in \ints \} 
\union \\
                   & \{ \npair{\stdcap[\noperm] } \} 
\union \\
                   & \{ \npair{\stdcap[readonly]} \mid \\
                   & \quad \exists r \in \RegionName \ldotp \\
                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)\} 
\union \\
                   & \{ \npair{\readwrite} \mid \\
                   & \quad \exists r \in \RegionName \ldotp \\
                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W) \land \\
                   & \qquad \qquad   \forall n' < n, W' \future W, v \ldotp \\
                   & \qquad \qquad \quad \heapSat{n'}{W'} \land \\
                   & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\Phi\update{a}{v}]{n''}{W''}
\union \\
                   & \{ \npair{\stdcap[\exec] = c} \mid \\
                   & \quad (\exists r \in \RegionName \ldotp \\
                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)) \\
                   & \quad \land \\
                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W') \ldotp \\
                   & \qquad \quad \npair[n']{[\pcreg \mapsto c, r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
\}
\end{align*}
\[
\stdkr \; : \;  \Worlds \monfun \UPred{\Words}%TODO: What should this be?
\]
\begin{align*}
  \stdkr(W) \defeq & \{ \npair{c} \mid \\
                   & \quad \forall v\ldotp \in \stdvr(W) \Rightarrow \\
                   & \qquad  \forall W' \future W\ldotp %on the whiteboard, we had public future world.
                     \forall \heapSat[\heap]{n'}{W'}, \npair{v_1 \dots v_k} \in \stdvr(W') \ldotp \\
                   & \qquad \quad ([\pcreg \mapsto \updatePcPerm{c},r_0 \mapsto v,r_1 \mapsto v_1,\dots, r_k \mapsto v_k], \heap) \diverge \}
\end{align*}

\[
\stder \; : \;  \Worlds \monfun \UPred{\Words}%TODO: What should this be?
\]
\begin{align*}
  \stder(W) \defeq & \{ \npair{\reg} \mid \\
                   & \quad \forall \npair{c} \in \stdkr(W) \\
                   & \qquad  \forall \heapSat{\heap}{n}{W} \ldotp \\
                   & \qquad \quad (\reg\update{r_0}{c},\heap) \diverge
\end{align*}

\begin{definition}[Heap staisfaction/erasure]
\begin{align*}
  \heapSat{n}{W} & & \text{iff} & &
                                \begin{gathered}
\exists \heapseg_{rem} \ldotp \Phi.\heap = \heapseg_{rem} \uplus \biguplus_{r\in \dom(W)} \heapseg_r\\
\text{and}\\
\forall r \in \dom(W) \ldotp \npair{\heapseg_r} \in W(r)(W)
\end{gathered}
\end{align*}
\end{definition}
\section{Related reading}
\label{sec:related-reading}

>>>>

\newcommand{\update}[2]{[#1 \mapsto #2]}
\newcommand{\sem}[1]{\left\llbracket #1 \right\rrbracket}

@@

\newcommand{\undefined}{\mathit{undefined}}

@@

%Memory projections
\newcommand{\plainproj}[1]{\mathrm{#1}}
\newcommand{\memheap}[1][\Phi]{#1.\plainproj{heap}}
\newcommand{\memreg}[1][\Phi]{#1.\plainproj{reg}}

\newcommand{\updateHeap}[3][\Phi]{#1\update{\plainproj{heap}.#2}{#3}}
\newcommand{\updateReg}[3][\Phi]{#1\update{\plainproj{reg}.#2}{#3}}

%Configuration end states
\newcommand{\failed}{\textsl{failed}}
\newcommand{\halted}{\textsl{halted}}

%Functions
\newcommand{\plainfun}[1]{\mathit{#1}}
\newcommand{\decode}{\plainfun{decode}}
\newcommand{\encode}{\plainfun{encode}}
\newcommand{\encodePerm}{\plainfun{encodePerm}}
\newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}(#1)}
\newcommand{\executeAllowed}[1]{\plainfun{executeAllowed}(#1)}
\newcommand{\nonZero}[1]{\plainfun{nonZero}(#1)}
\newcommand{\readAllowed}[1]{\plainfun{readAllowed}(#1)}
\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}(#1)}
\newcommand{\withinBounds}[1]{\plainfun{withinBounds}(#1)}

%World operations
\newcommand{\future}{\mathbin{\sqsupseteq}}
\newcommand{\heapSat}[3][\heap]{#1 :_{#2} #3}

@@

%LR
\newcommand{\intr}[2]{\mathcal{#1}\sem{#2}}
\newcommand{\valueintr}[1]{\intr{V}{#1}}
\newcommand{\exprintr}[1]{\intr{E}{#1}}
\newcommand{\contintr}[1]{\intr{K}{#1}}
\newcommand{\stdvr}{\valueintr{\asmType}}
\newcommand{\stder}{\exprintr{\asmType}}
\newcommand{\stdkr}{\contintr{\asmType}}
\newcommand{\observations}{\mathcal{O}}
\newcommand{\npair}[2][n]{\left(#1,#2 \right)}

@@

\begin{document}
$\RegName$ contains $\pcreg$, but is otherwise some undefined, finite set.
\begin{align*}
\Addrs &::= \nats & & &
\Words &::= \Caps + \ints \\
\Regs  &::= \RegName \rightarrow \Words & & &
\Heaps &::= \Addrs \rightarrow \Words \\
\Perms &::= \{ \noperm, \readonly, \readwrite, \exec, \entry, \rwx\} & & &
\Mems  &::= \Regs \times \Heaps \\
\Caps  &::= \Perms \times \Addrs \times \Addrs \times \Addrs & & &
\Confs &::= \Mems + \{\failed, \halted \times \Mems\}
\end{align*}

@@

  \path[every node/.style={font=\sffamily\small}]
    (1) edge (2)
    (2) edge (3)
    (3) edge (6)
    (1) edge (4)
    (4) edge (5)
    (5) edge (6);
\end{tikzpicture}
\caption{Permission hierarchy}
\label{fig:perm-hier}
\end{figure}
Notation:
$$\begin{array}{rcl}
i       &\in& \Instrs \\
r       &\in& \RegName\\
\mem    &::=& (\reg,\heap)\\
\pc     &\in& \Caps \\
\pcreg  &\in& \RegName \\
\Phi    &::=& \mem \in \Confs\\
\memheap&\in& \Heaps \\
\memreg &\in& \Regs \\
\addr   &\in& \Addrs\\
\perm   &\in& \Perms\\
(\perm,\start,\addrend,\addr) &\in& \Caps \\
n       &\in& \ints\\
\end{array}$$
Further definitions:
$$\begin{array}{rcl}
\lv    &::=& \refreg{r} \\
\hv    &::=& \refheap{r}\\
\rv    &::=& n \mid \lv \\
i      &::=& \fail \mid \halt \mid 
             \jmp{\lv} \mid \jnz{\lv}{\rv} \mid
             \isptr{\lv}{\rv} \mid \setptr{\lv}{\rv} \mid \\
       &   & \lea{\lv}{\rv} \mid\move{\lv}{\rv} \mid \load{\lv}{\hv} \mid \store{\hv}{\rv} \mid  \\
       &   & \restrict{\lv}{\rv}{\rv} \mid \subseg{\lv}{\rv}{\rv} \mid \plus{\lv}{\rv}{\rv}
\end{array}$$

\subsection*{Semantics}
Assume a $\decode$ function that decodes integer to instructions:

@@

\decode &:\ints \rightarrow \Instrs
Assume an $\encodePerm$ function that encodes a permission as an integer:
\encodePerm &: \Perms \rightarrow \ints
  \Phi & \rightarrow \sem{\decode(\memreg(\pcreg))}(\Phi) & &                                   
                                                              \arraycolsep=0pt
                                                                \text{if $\memreg(\pcreg) = \stdcap$}\\
                                                                \quad\text{and $\start \leq \addr < \addrend$}\\
                                                                \quad\text{and $\perm \in \{ \exec,\rwx \}$ }

@@

  \withinBounds{(\_,\start,\addrend,\addr)} &=
                                                \true  & \text{if $\start \leq \addr \leq \addrend$} \\

@@

@@

  \sem{\halt}(\Phi)                        & = (\halted,\Phi) \\
  \sem{\jmp{\lv}}(\Phi)                    & = 
                                               \updateReg{\pcreg}{\updatePcPerm{c}} &
                                               \arraycolsep=0pt
                                                 \text{if }\memreg(lv) = c \\
                                                 \quad\text{and }c=(\perm,\start,\addrend,\addr)\\
                                                 \quad\text{and }\executeAllowed{\perm}
                                               \end{array} \\

@@

  \sem{\jnz{\lv}{\rv}}(\Phi)               & = 
                                               \updateReg{\pcreg}{\updatePcPerm{\var{c}}} &
                                               \arraycolsep=0pt

@@

                                                 \text{if $\nonZero{\memreg(\rv)}$} \\ 
                                                 \quad\text{and $\memreg(lv) = c$} \\
                                                 \quad\text{and $c=(\perm,\start,\addrend,\addr)$}\\
                                                 \quad\text{and $\executeAllowed{\perm}$}

@@

                                               \updateReg{\pcreg}{\memreg(\pcreg) + 1} & \text{if not $\nonZero{\memreg(\rv)}$}\\

@@

                                               \updateReg{r_1}{\var{w}} &
                                               \arraycolsep=0pt

@@

                                                 \text{if } r_1 \neq \pcreg \\
                                                 \quad\text{and }\memreg(r_2) = (\perm,\start,\addrend,\addr) = \var{c} \\
                                                 \quad\text{and }\readAllowed{\perm} \text{ and } \withinBounds{\var{c}} \\
                                                 \quad\text{and }\var{w} = \memheap(\addr)

@@

                                               \updateHeap{\addr}{\var{w}} &
                                               \arraycolsep=0pt

@@

                                                 \text{if }\memreg(r_1) = (\perm,\start,\addrend,\addr) = \var{c} \\
                                                 \quad\text{and }\writeAllowed{\perm} \text{ and } \withinBounds{\var{c}} \\
                                                 \quad\text{and }\var{w} = \memreg(r_2)

@@

 \sem{\move{\refreg{r_1}}{\rv}}            & = \begin{cases}
                                                 \updateReg{r_1}{\memreg(\rv)} & \text{if $r_1 \neq \pcreg$} \\ % For now we do this, so you need to use a jump to move stuff into pc.
  \sem{\lea{\refreg{r_1}}{\rv}}            & =

@@

                                               \updateReg{r_1}{\var{c}} &
                                                 \arraycolsep=0pt
                                                   \text{if either $n = \rv$ or $\rv = \refheap{r_2}$ and $n = \memreg(r_2)$} \\
                                                   \quad\text{and in either case $n \in \ints $} \\
                                                   \quad\text{and $\memreg(r_1) = \stdcap$}\\
                                                   \quad\text{and $\var{c} = (\perm,\start,\addrend,\addr + n)$}
                                               \failed               & \text{otherwise}
\\ % In the M-Machine, lea checks whether the pointer stay within the allowed range in the pointer.
  \sem{\restrict{\refreg{r_1}}{\rv_1}{\rv_2}}           & =
                                               \updateReg{r_1}{\var{c}}  &
                                                 \arraycolsep=0pt
                                                   \text{if $r_1 \neq \pcreg$}\\
                                                   \quad\text{and $\memreg(\rv_1) = c'$}\\
                                                   \quad\text{and $c' = \stdcap$}\\
                                                   \quad\text{and either $\rv_2 = n$ or $\memreg(\rv_2) = n$}\\
                                                   \quad\text{and in either case $n \in \ints$}\\
                                                   \quad\text{and $\var{newPerm} = \encodePerm(n)$}\\
                                                   \quad\text{and $\var{newPerm} \sqsubseteq \perm$}\\
                                                   \quad\text{and $c = (\var{newPerm},\start,\addrend,\addr)$}
                                               \failed                   & \text{otherwise}
  \sem{\plus{\refreg{r_1}}{\rv_1}{\rv_2}}               & =
                                                            \updateReg{r_1}{n_1+n_2} &
                                                            \arraycolsep=0pt
                                                              \text{if $r_1 \neq \pcreg$}\\
                                                              \quad\text{and $r_1\neq \pcreg$}\\
                                                              \quad\text{for $i \in \{1,2\}$}\\
                                                              \qquad\text{$n_i = \rv_i$ or $n_i = \memreg{\rv_i}$}
                                                              \qquad\text{and in either case $n_i \in \ints$}

@@

  \sem{\isptr{\lv}{\rv}} & = \undefined \\ 
  \sem{\setptr{\lv}{\rv}} & = \undefined \\ 
  \sem{\subseg{\lv}{\rv}{\rv}} & = \undefined 

@@

\section{Examples}
\label{sec:examples}
\subsection{Ticket Dispenser}
\newcommand{\size}{\var{size}}
\label{sec:tick-disp}
\[
  \begin{array}{r l}
    \init: & \move{\refreg{r_1}}{\size} \\
           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
           & \lea{\refreg{r_0}}{3} \\
           & \jmp{\malloc} \\
           & \store{\refheap{r_1}}{(\encode(i_1))} \\
           & \lea{\refreg{r_1}}{1} \\
           & \store{\refheap{r_1}}{(\encode(i_2))} \\
           & \lea{\refreg{r_1}}{1} \\
           & \vdots \\
           & \store{\refheap{r_1}}{(\encode(i_6))} \\
           & \lea{\refreg{r_1}}{1} \\
           & \move{\refreg{r_0}}{\refreg{r_1}} \\
           & \lea{\refreg{r_0}}{1} \\
           & \store{\refheap{r_1}}{\refreg{r_0}} \\
           & \lea{\refreg{r_1}}{1} \\
           & \store{\refheap{r_1}}{0} \\
           & \lea{\refreg{r_1}}{-7} \\
           & \restrict{\refreg{r_1}}{(\encodePerm{(\entry)})} \\
           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
           & \lea{\refreg{r_0}}{4} \\
           & \restrict{\refreg{r_0}}{(\encodePerm{(\entry)})} \\
           & \jmp{\codelabel{adv}}
  \end{array}
\]
Here $\size$ is 9. $i_1 \dots i_7$ refers to the instructions in the following program:
\[
  \begin{array}{r l}
    i_1 :& \move{\refreg{r_2}}{\refreg{\pcreg}} \\
    i_2 :& \load{\refreg{r_2}}{\size - 2} \\
    i_3 :& \load{\refreg{r_1}}{\refheap{r_2}}\\
    i_4 :& \plus{\refreg{r_3}}{\refreg{r_1}}{2} \\
    i_5 :& \store{\refheap{r_2}}{\refreg{r_3}} \\
    i_6 :& \move{\refreg{r_2}}{0} \\
    i_7 :& \jmp{\refreg{r_0}}
  \end{array}
\]
The heap layout for the above program can be seen in Figure~\ref{tab:tick-disp-heap}.
\begin{figure}[h]
  \centering
  \begin{tabular}{ | c | l | }
    \hline
    Addr. & Content \\ \hline
    0   &  $i_1$ \\ \hline
    1   &  $i_2$ \\ \hline
    2   &  $i_3$ \\ \hline
    3   &  $i_4$ \\ \hline
    4   &  $i_5$ \\ \hline
    5   &  $i_6$ \\ \hline
    5   &  $i_7$ \\ \hline
    7   & Cap. for addr. 8 \\ \hline
    8   & The counter \\ \hline  
  \end{tabular}
  \caption{The heap layout for the ticket dispenser.}
  \label{tab:tick-disp-heap}
\end{figure}

\section{Logical Relation}
\label{sec:logical-relation}
\subsection{Worlds}
\[
\Worlds \cong \frac{1}{2} (\RegionName \parfun (\Worlds \monfun \UPred{\Addrs \parfun \Words}))
\]

\[
\stdvr \; : \;  \Worlds \monfun \UPred{\Words}
\]
\begin{align*}
  \stdvr(W) \defeq & \{ \npair{i} \mid i \in \ints \} 
\union \\
                   & \{ \npair{\stdcap[\noperm] } \} 
\union \\
                   & \{ \npair{\stdcap[\readonly]} \mid \\
                   & \quad \exists r \in \RegionName \ldotp \\
                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)\} 
\union \\
                   & \{ \npair{\readwrite} \mid \\
                   & \quad \exists r \in \RegionName \ldotp \\
                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W) \land \\
                   & \qquad \qquad   \forall n' < n, W' \future W, v \ldotp \\
                   & \qquad \qquad \quad \heapSat{n'}{W'} \land \\
                   & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \}
\union \\
                   & \{ \npair{\stdcap[\exec] = c} \mid \\
                   & \quad (\exists r \in \RegionName \ldotp \\
                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)) \\
                   & \quad \land \\
                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W') \ldotp \\
                   & \qquad \quad \npair[n']{[\pcreg \mapsto c, r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
\}
\end{align*}

Consider what the ``good observations'' should be? Before we do that - what should the end result of a computation be?
\begin{align*}
  \observations \quad : & \quad  \Worlds \monfun \UPred{\Confs} \\%TODO: What should this be?
  \observations (W) \defeq & \{ (\reg,\heap) \mid (\reg,\heap) \diverge \}
\end{align*}
\begin{align*}
  \stdkr \quad : & \quad  \Worlds \monfun \UPred{\Words} \\%TODO: What should this be?
  \stdkr(W) \defeq & \{ \npair{c} \mid \\
                   & \quad \forall v\ldotp \in \stdvr(W) \Rightarrow \\
                   & \qquad  \forall W' \future W\ldotp %on the whiteboard, we had public future world.
                     \forall \heapSat[\heap]{n'}{W'}, \npair{v_1 \dots v_k} \in \stdvr(W') \ldotp \\
                   & \qquad \quad ([\pcreg \mapsto \updatePcPerm{c},r_0 \mapsto v,r_1 \mapsto v_1,\dots, r_k \mapsto v_k], \heap) \in \observations(W) \}
\end{align*}

\begin{align*}
  \stder \quad : & \quad \Worlds \monfun \UPred{\Words} \\%TODO: What should this be?
  \stder(W) \defeq & \{ \npair{\reg} \mid \\
                   & \quad \forall \npair{c} \in \stdkr(W) \\
                   & \qquad  \forall \heapSat[\heap]{n}{W} \ldotp \\
                   & \qquad \quad (\reg\update{r_0}{c},\heap) \in \observations(W) \}
\end{align*}

\begin{definition}[Heap staisfaction/erasure]
\begin{align*}
  \heapSat{n}{W} & & \text{iff} & &
                                \begin{gathered}
\exists \heapseg_{rem} \ldotp \memheap = \heapseg_{rem} \uplus \biguplus_{r\in \dom(W)} \heapseg_r\\
\text{and}\\
\forall r \in \dom(W) \ldotp \npair{\heapseg_r} \in W(r)(W)
\end{gathered}
\end{align*}
\end{definition}
\section{Related reading}
\label{sec:related-reading}
